### 2019-01-28 write_up/C_changes.tex
Updated text and output latex table of results

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Mon Jan 28 16:59:10 2019
\begin{longtable}{rllrrrlllr}
\caption{Catch effects of the closure scenarios} \\ 
  \hline
scenario & metric & pop & before & after & diff & timescale & basis & data\_type & resolution \\ 
  \hline
 53 & Catch & spp\_3 & 112112.36 & 112521.76 & 0.37 & yearly & high\_pop & real\_pop & 20.00 \\ 
   49 & Catch & spp\_3 & 112112.36 & 113686.57 & 1.40 & yearly & high\_pop & survey & 20.00 \\ 
   31 & Catch & spp\_3 & 112112.36 & 114683.92 & 2.29 & yearly & high\_pop & commercial & 10.00 \\ 
   52 & Catch & spp\_3 & 112112.36 & 118512.95 & 5.71 & monthly & high\_pop & real\_pop & 20.00 \\ 
   51 & Catch & spp\_3 & 112112.36 & 120188.28 & 7.20 & weekly & high\_pop & real\_pop & 20.00 \\ 
   30 & Catch & spp\_3 & 112112.36 & 121325.90 & 8.22 & monthly & high\_pop & commercial & 10.00 \\ 
   29 & Catch & spp\_3 & 112112.36 & 122650.01 & 9.40 & weekly & high\_pop & commercial & 10.00 \\ 
   21 & Catch & spp\_3 & 112112.36 & 127010.38 & 13.29 & yearly & high\_pop & survey & 5.00 \\ 
   44 & Catch & spp\_3 & 112112.36 & 131095.21 & 16.93 & monthly & high\_pop & commercial & 20.00 \\ 
   35 & Catch & spp\_3 & 112112.36 & 133798.31 & 19.34 & yearly & high\_pop & survey & 10.00 \\ 
   38 & Catch & spp\_3 & 112112.36 & 135496.66 & 20.86 & monthly & high\_pop & real\_pop & 10.00 \\ 
    7 & Catch & spp\_3 & 112112.36 & 137987.16 & 23.08 & yearly & high\_pop & survey & 1.00 \\ 
   37 & Catch & spp\_3 & 112112.36 & 138717.85 & 23.73 & weekly & high\_pop & real\_pop & 10.00 \\ 
   43 & Catch & spp\_3 & 112112.36 & 139107.65 & 24.08 & weekly & high\_pop & commercial & 20.00 \\ 
   39 & Catch & spp\_3 & 112112.36 & 139763.03 & 24.66 & yearly & high\_pop & real\_pop & 10.00 \\ 
   15 & Catch & spp\_3 & 112112.36 & 143183.40 & 27.71 & weekly & high\_pop & commercial & 5.00 \\ 
    2 & Catch & spp\_3 & 112112.36 & 146551.20 & 30.72 & monthly & high\_pop & commercial & 1.00 \\ 
   45 & Catch & spp\_3 & 112112.36 & 146581.28 & 30.74 & yearly & high\_pop & commercial & 20.00 \\ 
   16 & Catch & spp\_3 & 112112.36 & 147719.91 & 31.76 & monthly & high\_pop & commercial & 5.00 \\ 
    1 & Catch & spp\_3 & 112112.36 & 149313.55 & 33.18 & weekly & high\_pop & commercial & 1.00 \\ 
   25 & Catch & spp\_3 & 112112.36 & 152756.42 & 36.25 & yearly & high\_pop & real\_pop & 5.00 \\ 
   23 & Catch & spp\_3 & 112112.36 & 155199.82 & 38.43 & weekly & high\_pop & real\_pop & 5.00 \\ 
   24 & Catch & spp\_3 & 112112.36 & 156523.19 & 39.61 & monthly & high\_pop & real\_pop & 5.00 \\ 
    3 & Catch & spp\_3 & 112112.36 & 157369.75 & 40.37 & yearly & high\_pop & commercial & 1.00 \\ 
   17 & Catch & spp\_3 & 112112.36 & 157786.23 & 40.74 & yearly & high\_pop & commercial & 5.00 \\ 
   10 & Catch & spp\_3 & 112112.36 & 160278.27 & 42.96 & monthly & high\_pop & real\_pop & 1.00 \\ 
    9 & Catch & spp\_3 & 112112.36 & 161276.13 & 43.85 & weekly & high\_pop & real\_pop & 1.00 \\ 
   11 & Catch & spp\_3 & 112112.36 & 165538.03 & 47.65 & yearly & high\_pop & real\_pop & 1.00 \\ 
   \hline
\hline
\label{tab:7}
\end{longtable}