### 2019-01-28 write_up/F_changes.tex

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Mon Jan 28 16:57:19 2019
\begin{longtable}{rllrrrlllr}
\caption{Fishing mortality effects of the closure scenarios} \\ 
  \hline
scenario & metric & pop & before & after & diff (\%) & timescale & basis & data\_type & resolution \\ 
  \hline
  9 & F & spp\_3 & 1.08 & 0.29 & -73.47 & weekly & high\_pop & real\_pop & 1.00 \\ 
   10 & F & spp\_3 & 1.08 & 0.29 & -72.94 & monthly & high\_pop & real\_pop & 1.00 \\ 
   11 & F & spp\_3 & 1.08 & 0.35 & -68.04 & yearly & high\_pop & real\_pop & 1.00 \\ 
   45 & F & spp\_3 & 1.08 & 0.58 & -46.70 & yearly & high\_pop & commercial & 20.00 \\ 
    1 & F & spp\_3 & 1.08 & 0.58 & -46.21 & weekly & high\_pop & commercial & 1.00 \\ 
   23 & F & spp\_3 & 1.08 & 0.59 & -45.27 & weekly & high\_pop & real\_pop & 5.00 \\ 
    2 & F & spp\_3 & 1.08 & 0.59 & -45.06 & monthly & high\_pop & commercial & 1.00 \\ 
    7 & F & spp\_3 & 1.08 & 0.60 & -44.48 & yearly & high\_pop & survey & 1.00 \\ 
   24 & F & spp\_3 & 1.08 & 0.61 & -43.20 & monthly & high\_pop & real\_pop & 5.00 \\ 
    3 & F & spp\_3 & 1.08 & 0.64 & -40.82 & yearly & high\_pop & commercial & 1.00 \\ 
   25 & F & spp\_3 & 1.08 & 0.65 & -39.94 & yearly & high\_pop & real\_pop & 5.00 \\ 
   17 & F & spp\_3 & 1.08 & 0.67 & -38.11 & yearly & high\_pop & commercial & 5.00 \\ 
   15 & F & spp\_3 & 1.08 & 0.71 & -34.38 & weekly & high\_pop & commercial & 5.00 \\ 
   43 & F & spp\_3 & 1.08 & 0.71 & -34.31 & weekly & high\_pop & commercial & 20.00 \\ 
   16 & F & spp\_3 & 1.08 & 0.73 & -32.58 & monthly & high\_pop & commercial & 5.00 \\ 
   51 & F & spp\_3 & 1.08 & 0.78 & -27.92 & weekly & high\_pop & real\_pop & 20.00 \\ 
   37 & F & spp\_3 & 1.08 & 0.78 & -27.76 & weekly & high\_pop & real\_pop & 10.00 \\ 
   39 & F & spp\_3 & 1.08 & 0.79 & -26.98 & yearly & high\_pop & real\_pop & 10.00 \\ 
   38 & F & spp\_3 & 1.08 & 0.81 & -25.47 & monthly & high\_pop & real\_pop & 10.00 \\ 
   21 & F & spp\_3 & 1.08 & 0.81 & -25.21 & yearly & high\_pop & survey & 5.00 \\ 
   35 & F & spp\_3 & 1.08 & 0.81 & -25.05 & yearly & high\_pop & survey & 10.00 \\ 
   44 & F & spp\_3 & 1.08 & 0.87 & -19.91 & monthly & high\_pop & commercial & 20.00 \\ 
   52 & F & spp\_3 & 1.08 & 0.88 & -18.39 & monthly & high\_pop & real\_pop & 20.00 \\ 
   30 & F & spp\_3 & 1.08 & 0.96 & -11.06 & monthly & high\_pop & commercial & 10.00 \\ 
   29 & F & spp\_3 & 1.08 & 0.98 & -9.80 & weekly & high\_pop & commercial & 10.00 \\ 
   31 & F & spp\_3 & 1.08 & 1.03 & -4.36 & yearly & high\_pop & commercial & 10.00 \\ 
   53 & F & spp\_3 & 1.08 & 1.06 & -1.64 & yearly & high\_pop & real\_pop & 20.00 \\ 
   49 & F & spp\_3 & 1.08 & 1.07 & -1.01 & yearly & high\_pop & survey & 20.00 \\ 
   \hline
\hline
\label{tab:6}
\end{longtable}