### 2019-02-19 write_up/F_changes_ratio.tex
Text updates and include test of effectiveness of knowledge parameters

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Tue Feb 19 17:01:38 2019
\begin{longtable}{rllrrrlllr}
\caption{Fishing mortality effects of the
     closure scenarios (based on highest ratio, ordered by most
effective first)} \\ 
  \hline
scenario & metric & pop & before & after & diff & timescale & basis & data\_type & resolution \\ 
  \hline
  6 & F & spp\_3 & 1.08 & 0.52 & -52.27 & yearly & high\_ratio & commercial & 1.00 \\ 
   48 & F & spp\_3 & 1.08 & 0.57 & -47.06 & yearly & high\_ratio & commercial & 20.00 \\ 
   50 & F & spp\_3 & 1.08 & 0.63 & -41.53 & yearly & high\_ratio & survey & 20.00 \\ 
   18 & F & spp\_3 & 1.08 & 0.71 & -34.23 & weekly & high\_ratio & commercial & 5.00 \\ 
   19 & F & spp\_3 & 1.08 & 0.72 & -33.42 & monthly & high\_ratio & commercial & 5.00 \\ 
   34 & F & spp\_3 & 1.08 & 0.78 & -27.75 & yearly & high\_ratio & commercial & 10.00 \\ 
    5 & F & spp\_3 & 1.08 & 0.80 & -25.99 & monthly & high\_ratio & commercial & 1.00 \\ 
   20 & F & spp\_3 & 1.08 & 0.81 & -25.27 & yearly & high\_ratio & commercial & 5.00 \\ 
    4 & F & spp\_3 & 1.08 & 0.85 & -21.52 & weekly & high\_ratio & commercial & 1.00 \\ 
   54 & F & spp\_3 & 1.08 & 0.89 & -17.46 & weekly & high\_ratio & real\_pop & 20.00 \\ 
   55 & F & spp\_3 & 1.08 & 0.89 & -17.46 & monthly & high\_ratio & real\_pop & 20.00 \\ 
   56 & F & spp\_3 & 1.08 & 0.89 & -17.46 & yearly & high\_ratio & real\_pop & 20.00 \\ 
   26 & F & spp\_3 & 1.08 & 0.92 & -14.73 & weekly & high\_ratio & real\_pop & 5.00 \\ 
   27 & F & spp\_3 & 1.08 & 0.92 & -14.73 & monthly & high\_ratio & real\_pop & 5.00 \\ 
   28 & F & spp\_3 & 1.08 & 0.92 & -14.73 & yearly & high\_ratio & real\_pop & 5.00 \\ 
   13 & F & spp\_3 & 1.08 & 0.96 & -11.53 & monthly & high\_ratio & real\_pop & 1.00 \\ 
   14 & F & spp\_3 & 1.08 & 0.96 & -11.01 & yearly & high\_ratio & real\_pop & 1.00 \\ 
   12 & F & spp\_3 & 1.08 & 0.97 & -10.66 & weekly & high\_ratio & real\_pop & 1.00 \\ 
   32 & F & spp\_3 & 1.08 & 1.02 & -5.94 & weekly & high\_ratio & commercial & 10.00 \\ 
   22 & F & spp\_3 & 1.08 & 1.02 & -5.64 & yearly & high\_ratio & survey & 5.00 \\ 
   33 & F & spp\_3 & 1.08 & 1.02 & -5.29 & monthly & high\_ratio & commercial & 10.00 \\ 
   36 & F & spp\_3 & 1.08 & 1.03 & -4.52 & yearly & high\_ratio & survey & 10.00 \\ 
   40 & F & spp\_3 & 1.08 & 1.03 & -4.52 & weekly & high\_ratio & real\_pop & 10.00 \\ 
   41 & F & spp\_3 & 1.08 & 1.03 & -4.52 & monthly & high\_ratio & real\_pop & 10.00 \\ 
   42 & F & spp\_3 & 1.08 & 1.03 & -4.52 & yearly & high\_ratio & real\_pop & 10.00 \\ 
   46 & F & spp\_3 & 1.08 & 1.04 & -3.50 & weekly & high\_ratio & commercial & 20.00 \\ 
    8 & F & spp\_3 & 1.08 & 1.06 & -2.42 & yearly & high\_ratio & survey & 1.00 \\ 
   47 & F & spp\_3 & 1.08 & 1.09 & 0.52 & monthly & high\_ratio & commercial & 20.00 \\ 
   \hline
\hline
\label{tab:7}
\end{longtable}