### 2019-09-17 write_up/F_changes.tex
All revisions to figures

% latex table generated in R 3.6.1 by xtable 1.8-4 package
% Tue Sep 17 17:32:36 2019
\begin{longtable}{rlrrrllr}
\caption{Fishing mortality effects of the
	     closure scenarios (ordered by most effective first)} \\ 
  \hline
Scenario No & Population & F before & F after & \% F change & data type & timescale & resolution \\ 
  \hline
  9 & Population 3 & 1.08 & 0.29 & -73.47 & True Population & weekly & 1.00 \\ 
   10 & Population 3 & 1.08 & 0.29 & -72.94 & True Population & monthly & 1.00 \\ 
   11 & Population 3 & 1.08 & 0.35 & -68.04 & True Population & yearly & 1.00 \\ 
   45 & Population 3 & 1.08 & 0.58 & -46.70 & commercial & yearly & 20.00 \\ 
    1 & Population 3 & 1.08 & 0.58 & -46.21 & commercial & weekly & 1.00 \\ 
   23 & Population 3 & 1.08 & 0.59 & -45.27 & True Population & weekly & 5.00 \\ 
    2 & Population 3 & 1.08 & 0.59 & -45.06 & commercial & monthly & 1.00 \\ 
    7 & Population 3 & 1.08 & 0.60 & -44.48 & survey & yearly & 1.00 \\ 
   24 & Population 3 & 1.08 & 0.61 & -43.20 & True Population & monthly & 5.00 \\ 
    3 & Population 3 & 1.08 & 0.64 & -40.82 & commercial & yearly & 1.00 \\ 
   25 & Population 3 & 1.08 & 0.65 & -39.94 & True Population & yearly & 5.00 \\ 
   17 & Population 3 & 1.08 & 0.67 & -38.11 & commercial & yearly & 5.00 \\ 
   15 & Population 3 & 1.08 & 0.71 & -34.38 & commercial & weekly & 5.00 \\ 
   43 & Population 3 & 1.08 & 0.71 & -34.31 & commercial & weekly & 20.00 \\ 
   16 & Population 3 & 1.08 & 0.73 & -32.58 & commercial & monthly & 5.00 \\ 
   51 & Population 3 & 1.08 & 0.78 & -27.92 & True Population & weekly & 20.00 \\ 
   37 & Population 3 & 1.08 & 0.78 & -27.76 & True Population & weekly & 10.00 \\ 
   39 & Population 3 & 1.08 & 0.79 & -26.98 & True Population & yearly & 10.00 \\ 
   38 & Population 3 & 1.08 & 0.81 & -25.47 & True Population & monthly & 10.00 \\ 
   21 & Population 3 & 1.08 & 0.81 & -25.21 & survey & yearly & 5.00 \\ 
   35 & Population 3 & 1.08 & 0.81 & -25.05 & survey & yearly & 10.00 \\ 
   44 & Population 3 & 1.08 & 0.87 & -19.91 & commercial & monthly & 20.00 \\ 
   52 & Population 3 & 1.08 & 0.88 & -18.39 & True Population & monthly & 20.00 \\ 
   30 & Population 3 & 1.08 & 0.96 & -11.06 & commercial & monthly & 10.00 \\ 
   29 & Population 3 & 1.08 & 0.98 & -9.80 & commercial & weekly & 10.00 \\ 
   31 & Population 3 & 1.08 & 1.03 & -4.36 & commercial & yearly & 10.00 \\ 
   53 & Population 3 & 1.08 & 1.06 & -1.64 & True Population & yearly & 20.00 \\ 
   49 & Population 3 & 1.08 & 1.07 & -1.01 & survey & yearly & 20.00 \\ 
   \hline
\hline
\label{tab:6}
\end{longtable}