### 2019-09-24 write_up/F_changes.tex
JJP updates

% latex table generated in R 3.6.1 by xtable 1.8-4 package
% Tue Sep 24 20:53:17 2019
\begin{longtable}{rrrllr}
\caption{Fishing mortality effects of the
     closure scenarios on population 3 (ordered by most effective
first). The fishing mortality
     rate before the closure was 1.08.} \\ 
  \hline
Scenario No & F after closure & \% F change & data type & timescale & resolution \\ 
  \hline
  9 & 0.29 & -73.47 & true Population & weekly & 1.00 \\ 
   10 & 0.29 & -72.94 & true Population & monthly & 1.00 \\ 
   11 & 0.35 & -68.04 & true Population & yearly & 1.00 \\ 
   45 & 0.58 & -46.70 & commercial & yearly & 20.00 \\ 
    1 & 0.58 & -46.21 & commercial & weekly & 1.00 \\ 
   23 & 0.59 & -45.27 & true Population & weekly & 5.00 \\ 
    2 & 0.59 & -45.06 & commercial & monthly & 1.00 \\ 
    7 & 0.60 & -44.48 & survey & yearly & 1.00 \\ 
   24 & 0.61 & -43.20 & true Population & monthly & 5.00 \\ 
    3 & 0.64 & -40.82 & commercial & yearly & 1.00 \\ 
   25 & 0.65 & -39.94 & true Population & yearly & 5.00 \\ 
   17 & 0.67 & -38.11 & commercial & yearly & 5.00 \\ 
   15 & 0.71 & -34.38 & commercial & weekly & 5.00 \\ 
   43 & 0.71 & -34.31 & commercial & weekly & 20.00 \\ 
   16 & 0.73 & -32.58 & commercial & monthly & 5.00 \\ 
   51 & 0.78 & -27.92 & true Population & weekly & 20.00 \\ 
   37 & 0.78 & -27.76 & true Population & weekly & 10.00 \\ 
   39 & 0.79 & -26.98 & true Population & yearly & 10.00 \\ 
   38 & 0.81 & -25.47 & true Population & monthly & 10.00 \\ 
   21 & 0.81 & -25.21 & survey & yearly & 5.00 \\ 
   35 & 0.81 & -25.05 & survey & yearly & 10.00 \\ 
   44 & 0.87 & -19.91 & commercial & monthly & 20.00 \\ 
   52 & 0.88 & -18.39 & true Population & monthly & 20.00 \\ 
   30 & 0.96 & -11.06 & commercial & monthly & 10.00 \\ 
   29 & 0.98 & -9.80 & commercial & weekly & 10.00 \\ 
   31 & 1.03 & -4.36 & commercial & yearly & 10.00 \\ 
   53 & 1.06 & -1.64 & true Population & yearly & 20.00 \\ 
   49 & 1.07 & -1.01 & survey & yearly & 20.00 \\ 
   \hline
\hline
\label{tab:6}
\end{longtable}