### 2019-10-11 write_up/cover_letter.tex
Adding cover letter and highlights

\documentclass[12pt]{letter}
\usepackage[margin=2.54cm, a4paper]{geometry}
\signature{\vspace*{-1cm} 
	Paul Dolder \\ 
	Dr C\'oil\'in Minto\\ 
	Prof Jean-Marc Guarini \\
	Dr Jan Jaap Poos}
\address{Paul Dolder \\
	Marine and Freshwater Research Centre \\ 
	Galway-Mayo Institute of Technology \\ 
	Dublin Road \\ 
	Galway \\ 
	Ireland}

\begin{document}

\begin{letter}{Prof. Brian D. Fath \\ Department of Biological Sciences \\
		Towson University \\ 8000 York Road \\ Towson \\ Maryland 21252 \\ USA}
\opening{Dear Professor Fath,}

We would like to submit an \emph{Original research} paper titled \emph{`Highly
	resolved spatiotemporal simulations for exploring mixed fisheries
	dynamics'} for consideration in \emph{Ecological modelling}. In this
paper we develop and apply a simulation model of a ecological-fishery system
where individual fishing vessels exploit multiple heterogeneously distributed
fish stocks with full spatiotemporal population dynamics. In it we explore the
importance of spatial and temporal scale when modelling fishery interactions,
particularly in relation to modelling observational data from a system with
preferential sampling.\\

Our study tackles the issue of how fishers exploit a dynamic natural resource
with uncertain knowledge of its distribution. We detail how this affects our
understanding of the fisheries interactions with multiple fish populations,
which cannot be achieved with a conventional modelling approach. By capturing
dynamics observed in fisheries data in a novel simulation model where the
entire dynamics of the system (including the true spatial distribution of the
populations) are known, we expose that degrading the spatial and temporal
resolution of data through aggregation reduces our ability to define effective
spatial management measures. This would not of been possible without the highly
details and resolved simulation approach used here. The model has many
potential additional applications which we discuss, with the model made
available as an R software package for users to explore.\\

Should you consider our work suitable for review, we would like to suggest the
following associate editor and potential reviewers with a rationale below each:

\begin{itemize} 
	\item {\bf Associate Editor} Prof. Terrance Quinn II, College of
		Fisheries and Oceans, University of Fairbanks, Alaska
		(tjquinnii@alaska.edu).\\ 
	\item {\bf Reviewer} Dr David Peel, Commonwealth Scientific and
		Industrial Research Organisation (CSIRO), Castray Esplanade,
		Hobart TAS 7000, Australia (david.peel@csiro.au). \\
		Established expert in finite mixture modelling.  
	\item {\bf Reviewer} Dr John Morrengiello, University of Melbourne,
		Melbourne, Australia (john.morrongiello@unimelb.edu.au) \\
		Early career scientist with expertise in quantitative aquatic
		ecology and evolution along with publications on growth
		modelling.  
	\item {\bf Reviewer} Dr Ken Newman, U.S. Fish and Wildlife Service,
		Lodi Fish and Wildlife Office, 850 Guild Avenue, Lodi, CA
		95240, USA (Ken\_Newman@fws.gov)\\ Established expert in
		mathematical and applied statistics with application to natural
		resource management.  
\end{itemize}

Thank you for your time and consideration.

\closing{Yours Sincerely,} \end{letter}

\end{document}