### 2019-10-11 write_up/highlights.tex

\documentclass{article}

\begin{document}

\begin{itemize}

	\item one
	\item two
	\item three
	\item four
	\item five 
	\item six

\end{itemize}

\end{document}