### 2019-10-13 write_up/CreDit.tex
Final steps before submission

\documentclass{article}

\usepackage[utf8]{inputenc}
\begin{document}

CRediT author statement:

\begin{itemize}
	\item \textbf{Paul Dolder:} Conceptualisation, methodology, software,
		writing - original draft. 
	\item \textbf{Cóilín Minto:} Conceptualisation, Methodology, writing - review
	and editing, supervision, funding acquisition. 
	\item \textbf{Jean-Marc Guarini:} methodology, writing - review and
		editing, funding acquisition, supervision. 
	\item \textbf{Jan Jaap Poos:} methodology, writing - review and
		editing, supervision. \

\end{itemize}

\end{document}