### 2019-10-13 write_up/highlights.tex

\documentclass{article}

\begin{document}

\begin{itemize}

	\item We developed a highly spatially resolved discrete-event
		simulation model of mixed fishery interactions with multiple
		populations.	
	\item The simulation framework incorporates i) delay-difference population
		dynamics, ii) population movement using Gaussian Markov Fields
		to simulate patchy, heterogeneously distribution populations,
		and iii) individual-based fishery dynamics for multiple fleet
		characteristics using an explore-exploit strategy.
	\item The framework allows users to explore assumptions in modelling
		observational data and examine dynamics at a fine
		spatiotemporal scale. 
	\item We simulate fifty years of multiple fisheries exploiting four
		separate populations with different demographics and find
		nuanced spatio-temporal patterns in exploitation.
	\item A simulated spatial closure shows that when aggregating point
		data on fishing activity to a grid, the scale of aggregation
		matters for the ability to achieve fishery-conservation
		objectives.
	\item The simulation framework is available as an R package and we
		suggest multiple potential uses of the package (e.g. survey
		design evaluation, testing index standardisation, in-year
		fishery and biological modelling) for researchers.

\end{itemize}

\end{document}