### 2020-01-26 write_up/mixfishsim_ecolmod_revisions.tex
Incorparting Coilin and Jean-Marc's comments

\documentclass[review]{elsarticle}

\usepackage{xr} %% For cross-referencing supl figures
\externaldocument[S-]{Supplementary}

\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathtools}

%% for the annex figure
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,chains,calc}
% Set up a few colours for tikz figure
\colorlet{lcfree}{green}
\colorlet{lcnorm}{blue}
\colorlet{lccong}{red}

\usepackage{hhline}
\usepackage{booktabs}
\usepackage{eurosym}
\usepackage{longtable}

\usepackage{enumitem}
\usepackage{comment}

\usepackage[final]{changes}

\modulolinenumbers[1]

\journal{Ecological Modelling}

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%%%%%%%%%%%%%%%%%%%%%%%
%% correct equation line numbering

\let\oldequation\equation
\let\oldendequation\endequation
\renewenvironment{equation}
 {\linenomathNonumbers\oldequation}
 {\oldendequation\endlinenomath}

%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}
\title{Highly resolved spatiotemporal simulations for
	exploring mixed fishery dynamics}

%% Group authors per affiliation:
\author[1,2]{Paul J. Dolder\corref{c}}
\cortext[c]{Corresponding author}
\ead{paul.dolder@cefas.co.uk}

\author[1]{Cóilín Minto}
\author[3]{Jean-Marc Guarini}
\author[4,5]{Jan Jaap Poos}

\address[1]{Marine and Freshwater Research Centre, Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{Sorbonne Université, Faculty of Sciences, 4 Place Jussieu, 75005
	Paris, France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}
\address[5]{Aquaculture and Fisheries Group, Wageningen University \& Research,
	Zodiac Building 122, De Elst 1, 6708 WD Wageningen, the Netherlands}

\begin{abstract}

%%
To understand how data resolution impacts inference on mixed fisheries
interactions we developed a highly resolved spatiotemporal discrete-event
simulation model \emph{MixFishSim} incorporating: i) delay-difference
population dynamics, ii) population movement using Gaussian Random Fields to
simulate patchy, heterogeneously distributed \added{and moving fish}
populations, and iii) fishery dynamics for multiple fleet characteristics based
on population targeting under an explore-exploit strategy.\deleted{ This is
	implemented via a mix of correlated random walk movement (for
	exploration) and learned behaviour (for exploitation) phases of the
	fisheries. }\\ 
%%
\deleted{Fifty years of sub-daily fishing activity was simulated and used to
	draw inference on the underlying community structures. W}\added{We
	applied \emph{MixFishSim} to\deleted{draw}}\deleted{compared}
infer\deleted{ence}\deleted{s} \added{\deleted{on}community structure}
\added{when using data generated from:} \deleted{based on}commercial catch,
a\deleted{simulated} fixed-site sampling survey design and the true
\added{(simulated)} underlying populations.  \deleted{W}\deleted{\added{In
		doing so w}}\added{W}e \added{thereby} \deleted{i)}establish
the potential limitations of fishery-dependent data in providing a robust
\replaced{characterisation}{picture} of spatiotemporal distributions.\deleted{;
	and then ii) simulated an area closure based on areas defined from the
	different data sources at a range of temporal and spatial
	resolutions.}\deleted{Our framework allows users to explore the
	assumptions in modelling observational data and evaluate the underlying
	dynamics of such approaches at a fine spatial and temporal scale. In
	application to a mixed fishery exploiting four different populations w}
\deleted{\added{W}e found d}\added{D}ifferent spatial patterns were evident and
\deleted{that }the effectiveness of the spatial closure reduced when data were
aggregated across larger spatial areas. \deleted{; and then
	ii)}\deleted{\added{We then simulated an area closure based on areas
		defined from the different data sources at a range of temporal
		and spatial resolutions.}} \added{A simulated area
	closure showed that} \replaced{a}{A}ggregation across time periods has
less of a negative impact on the closure success than \added{aggregation} over
space. \deleted{ and w}\added{W}hile not as effective as when based on on the
true population, closures based on high catch rates observed in commercial data
were still able to reduce fishing on a protected species. \\

\added{Our framework allows users to explore the assumptions in modelling
	observational data and evaluate the underlying dynamics of such
	approaches at a fine spatial and temporal resolutions. From our
	application w}\deleted{W}e conclude that commercial data, while
containing bias, provide a useful tool for managing catches in mixed fisheries
if applied at the correct spatiotemporal scale. \\
%%
\end{abstract}

\begin{keyword}
spatiotemporal \sep mixed fisheries \sep individual based \sep spatial
management \sep heterogeneity \sep bycatch avoidance 
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions.
As fishers do not have full control over what species they select when fishing
in `mixed fisheries' it can result in catch of low quota or protected species.
If over-quota catch of a species for which they have no quota is discarded
without being accounted for it limits our ability to control fishing mortality
\citep{Alverson1994, Crowder1998, Rijnsdorp2007} and the ability to manage
fisheries for the biological and economic sustainability \citep{Ulrich2011a,
	Batsleer2015}.\\

There is increasing interest in technical solutions such as gear and spatial
closures as measures to reduce unwanted catch\added{ which is often discarded}
\citep{Kennelly2002, Catchpole2008, Bellido2011, Cosgrove2019}\added{.
	A}\deleted{ and a}daptive spatial management strategies have been
proposed as a way of reducing over-quota discards \citep{Holmes2011,
	Little2014, Dunn2014a}. However, if fisheries are to reduce unwanted
catch through spatial avoidance, an in-depth understanding of spatiotemporal
fishery dynamics is required. \\

Understanding the correct scale for spatial management measures to be effective
is crucial as it enables implementation of effective solutions which minimise
economic impact \citep{Dunn2016}. For example, the problem can be to identify a
scale that promotes species avoidance for vulnerable or low quota species while
allowing continuance of sustainable fisheries for available quota species.
Identifying the correct spatial scale remains a challenge because data on fish
location at high temporal and spatial resolutions is expensive and difficult to
collect and proxies are usually inferred from scientific surveys or commercial
catches with limited spatial and temporal resolution. Thus, implementation of
spatial measures is hampered by a lack of knowledge of fish and fishery
spatiotemporal dynamics and understanding of the scale at which these processes
become important for management. \\ \\

Identifying appropriate spatial scales for fisheries closures \deleted{has been a
highlighted a}\added{i}s crucial to their success \citep{Costello2010, Dunn2016}.
Inference on fisheries spatial dynamics is hampered where spatial information
is coarse due to low resolution reporting of fisheries catch which is
aggregated across larger gridded areas \citep{Branch2005}. Further, if data
does not allow identification of spatial features it may lead to poorly sited
closures \replaced{that}{which} are ineffectual or have unintended consequences. For example,
increased benthic impact on previously unexploited areas from the cod closure
in the North Sea were observed without the intended effect of reducing cod
exploitation \citep{Rijnsdorp2001,Dinmore2003}).\\

More refined spatiotemporal information has since become available through the
combination of logbook and Vessel Monitoring System (VMS) data \citep{Lee2010,
	Bastardie2010, Gerritsen2012, Mateo2016} and more real-time spatial
management has been possible \citep[e.g.][]{Holmes2011}.  However, fishers
establish favoured fishing grounds through an explore-exploit strategy
\citep{Rijnsdorp2011, Bailey2018} where they search for areas with high catches
and then use experience to return to areas where they have experienced high catch
in the past. This leads to an inherently biased sampling where target species
are over-represented in the catch as fishers exploit areas of high abundance.
There is a need to understand the influence of these biases on any spatial
management measures which are implemented based on inference from commercial
landings or catch data. \\

To understand the effect of spatiotemporal aggregation of data and fishery
targeting on our perception of spatial abundance of different fish populations
we ask two fundamental questions regarding inference derived from observational
data:

\begin{enumerate}
	\item Do different source\added{s} of sampling-derived fisheries data
		reflect\deleted{s} the underlying community structure? 	
	\item How do\deleted{es} data aggregation and data source impact on the success
		of spatial fisheries management measures?
\end{enumerate}

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time, using a Gaussian spatial
process to define suitable habitat for different populations. As the precise
locations of the fish are known directly rather than inferred from sampling or
commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries-independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries-independent
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

We use these evaluations to draw inference on the utility of commercial data in
supporting management decisions.

\section{Materials and Methods}

A Discrete-event simulation (DES) model of a hypothetical fishery was developed
as a software package (\textit{MixFishSim}). The modular approach enabled
efficient computation by allowing for sub-modules implemented on time-scales
appropriate to capture \deleted{the }characteristic\added{s} of the different processes (Figure
1). \replaced{S}{The following s}ub-modules \deleted{were included }to capture the full system \added{comprised}: 1)
\replaced{p}{P}opulation dynamics, 2) \replaced{r}{R}ecruitment dynamics, 3) \replaced{p}{P}opulation movement, 4)
fishery dynamics.\\

Population dynamics \added{for any number of species, as chosen by the user,}
operate on a daily time-step \added{(with recruitment occurring only during
	defined seasons for each population)}, while population movement occurs
on a weekly time-step, with the fishing module operating on a tow-by-tow basis
(i.e.\added{,} multiple events a day). \\

\subsection{Population dynamics}

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model \replaced{that}{which} models the fish populations in
terms of aggregate biomass of recruits and mature components rather than
keeping track of individuals \citep{Deriso1980, Schnute1985, Dichmont2003}. A
daily time-step was chosen to discretise continuous population processes on a
biologically relevant and computationally tractable timescale. Population
biomass growth was modelled as a function of previous recruited biomass,
intrinsic population growth and recruitment functionally linked to the adult
population size.  Biomass for each cell $c$ was incremented each day $d$ as
follows (\replaced{see}{the full parameter list is detailed in } Table \ref{tab: pop_variable} \added{for all parameter details}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& \left(1 + \rho \right) B_{c,d} \cdot \mathrm{e}^{-Z_{c,d}} - \rho \cdot
	\mathrm{e}^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& \left(B_{c,d-1} \cdot \mathrm{e}^{-Z_{c,d-1}} + W_{R-1} \cdot \left( \alpha_{d-1} \cdot
	R_{\tilde{y}\left(c\right)}\right) \right)
	\hspace{0.9cm} + \\
	& W_{R} \cdot \left( \alpha_{d} \cdot R_{\tilde{y} \left(c\right)}
	\right) \\
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient\textcolor{blue}{(Check)}, shown to be equal
to $\mathrm{e}^{-K}$ when $K$ is the \deleted{growth }rate \added{at which the
	asymptote is approached} from a von Bertalanffy \deleted{logistic
}growth model \citep{Schnute1985}. $W_{R-1}$ is the average weight of fish
prior to recruitment, while $W_{R}$ is the average recruited weight.
$\alpha_{d}$ represents the proportion of fish recruited during that day for
the year, while $R_{c,\tilde{y}(c)}$ is the annual recruits in year $y$ for
cell $c$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch is the sum of
those across all fleets and vessels, $C_{c, d} =
\sum\limits_{fl=1}^{FL}\sum\limits_{v=1}^{V_{fl}} E_{fl, v, c, d} \cdot Q_{fl}
\cdot D_{c, d}$ with $fl$ and $FL$ the fleet and total number of fleets, $v$
and $V_{fl}$ the vessel and total number of vessels per fleet respectively and
$E_{fl,v,c,d}$ and $Q_{fl}$ fishing effort and catchability of the gear, and
$D_{c,d}$ is the density of the population at the location fished. \\

\subsection{Recruitment dynamics}

Recruitment is modelled as a function of adult biomass. In
\textit{MixFishSim}, it can either take the form of a stochastic Beverton-Holt
stock recruitment relationship, or a stochastic Ricker stock recruitment
relationship. The Beverton-Holt relationship is defined as\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \deleted{\cdot} S_{c,d})}{(\beta + S_{c,d})} \\
	     \added{\ln(}R_{c,d}\added{)} \sim & \deleted{\log} N[(\replaced{\ln}{\log}(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes. The stochastic Ricker form \citep{Ricker1954} is:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
	\ln(R_{c,d}) \sim & N[(\ln(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the
density-dependent reduction in productivity as the SSB increases.

\subsection{Population movement dynamics}

Population movement is a combination of directed (advective) movement where at
certain times of year the population moves towards spawning grounds by
increasing the probabilities of moving into the spawning grounds from adjacent
cells, and random (diffusive) movement, governed by a stochastic process where
movement between adjacent cells is described by a set of probabilities.
\added{Stochastic probabilities are affected by the suitability of habitat,
	temperature in a cell and the thermal tolerance of a population to that
	temperature}.\\

\added{The combined process results in a population structure and movement
	pattern unique to each population, with population movement occurring
	on a weekly basis. \replaced{M}{The decision to m}odel\added{ing}
	population movement on a weekly timescale\deleted{ was to}
	reflect\added{s} that fish tend to aggregate in
	species\added{-}specific locations\deleted{ that have been} observed to
	last \replaced{between}{around } one \replaced{and}{to} two weeks
	\citep{Poos2007}. Therefore this process approximated the demographic
	shifts in fish populations throughout a year with seasonal spawning
	patterns }(Figure \ref{S-fig:5}).

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. \added{We use the most commonly used
	Matérn covariance structure as it is a flexible form that \replaced{contains the}{under certain
	conditions is of the same form as an} exponential \added{and double exponential as special cases}\deleted{ function} and
	it}\deleted{This} enables us to model the spatial autocorrelation
observed in animal populations where density is more similar in nearby
locations\deleted{, \added{but that correlation decreases non-linearly}}
\citep{Tobler1970, F.Dormann2007, Poos2007}. We change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field with an anisotropic pattern (to simulate a depth gradient) and
combine it with a temporally dynamic thermal tolerance field to imitate two key
drivers of population dynamics without modelling the processes explicitly. Each
population was initialised at a single location, and subsequently moved across
the entire space according to a probabilistic distribution based on habitat
suitability (represented by the normalised values from the GRFs), temperature
tolerance and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}$ is the index of habitat
suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see below).
\\

During pre-defined weeks of the year the habitat suitability is modified with
user-defined spawning habitat locations, resulting in each population having
concentrated areas where spawning takes place. The populations then move
towards these cells in the weeks prior to spawning, resulting in directional
movement towards the spawning grounds. \\

A time-varying temperature covariate changes the \deleted{interaction between time and}
suitable habitat on a weekly time-step. Each population $p$ was assigned a
thermal tolerance with mean, $\mu_{p}$ and standard deviation, $\sigma_{p}$ so that
each cell and population temperature tolerance is defined \replaced{as}{that}:
\begin{equation}
	Tol_{c, p, wk} = \frac{1}{ \sqrt{2\pi  \sigma^2_{p}}}  \exp \left(-
		\frac{(T_{c, wk} - \mu_{p})^2}{2  \sigma^2_{p}} \right)	
\end{equation}
Where $Tol_{c, p, wk}$ is the tolerance of population $p$ for cell $c$ in week
$wk$, $T_{c, wk}$ is the temperature in the cell given the week and $\mu_{p}$
and $\sigma_{p}$ the mean and standard deviation of the population temperature
tolerance \replaced{(see Table \ref{tab:pop_move} for variable descriptions).}{. The variables, their meaning and units for population movement is
provided in Table \ref{tab:pop_move}}. \\

\deleted{The final combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks (Poos et al 2007). Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).}

\subsection{Fleet dynamics}

Fleet dynamics \replaced{were}{can be }broadly categorised into three
components. \textit{Fleet targeting} \replaced{determined}{determines} the
fleet catch efficiency and preference towards a particular population;
\textit{trip-level decisions} \replaced{determined}{that determines} the
initial location to be fished at the beginning of a trip; and
\textit{within-trip decisions} \replaced{determined fishing locations}{, that
	determines movement from one fishing spot to another} within a trip.
\replaced{This results in an explore-exploit strategy}{ was implemented in the
	model that combined these three components} for individual vessels to
maximise their catch from an unknown resource distribution \citep{Bailey2018}.
The decision to use an individual based model for fishing vessels was taken
because fishers are heterogeneous in their location choice behaviour due to
different objectives, risk preference and targeting preference
\citep{VanPutten2012a, Boonstra2015}.  Therefore fleet dynamics are emergent
from individual dynamics rather than pre-defined group dynamics.

\subsubsection{Fleet targeting}

Each fleet of \textit{$n_{fl}$} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$ which
are each bound by [0,1]. The product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affects the overall catch rates for the fleet and the preferential
targeting of one species over another. This, in combination with the parameter
choice for the step-function defined below (as well as some randomness from the
exploratory fishing process) determined the preference of fishing locations for
the fleet.

\subsubsection{Decision about where to fish at the start of a trip}

Several studies \citep[for a review see][]{Girardin2016} have confirmed past activity
and past catch rates are strong predictors of fishing location choice. For this
reason, the fleet dynamics sub-model included a learning component, where a
vessel's initial fishing location in a trip was based on selecting from
previously successful fishing locations. This was achieved by calculating an
expected revenue based on the catches from locations fished in the preceding
trip as well as the same month periods in previous years and the travel costs
from the port to the fishing grounds. Then a vessel chooses randomly from the
top 70 \% of fishing events (defined as the `threshold') in terms of expected
profit within that season.

\subsubsection{Decision about where to fish within a trip}

Fishing locations within a trip are initially determined by a modified random
walk process. As the simulation progresses the within-trip decision become
gradually more influenced by experience gained from past fishing locations (as
per the initial trip-level location choice), moving location choice towards
areas of higher perceived profit. A random walk was chosen for the exploratory
fishing process as it is the simplest assumption commonly used in ecology to
describe optimal animal search strategy for exploiting heterogeneously
distributed prey about which there is uncertain knowledge
\citep{Viswanathan1999}. In a random walk, movement is a stochastic process
through a series of steps. These steps have a length, and a direction that can
either be equal in length or take some other functional form. The direction of
the random walk was also correlated (known as `persistence') providing some
overall directional movement \citep{Codling2008}. \\

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: (\textcolor{blue}{[CM: use $\ln$ where we use the natural logarithm. Double-check this equation is right as per the code. As it looks different here: \url{https://github.com/pdolder/MixFishSim/blob/master/R/step_length.R]}. There it looks to simplify to $b_2(b_1/b_2)^{Rev/b_3}$, assuming parameters are positive.})
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
($x_{t}$,$y_{t}$) to ($x_{t+1}$, $y_{t+1}$) is defined by:
\begin{equation}
	\begin{split}
 (x_{t+1}, y_{t+1}) =  & x_{t} + Le \cdot \cos \left(\frac{\pi \cdot
		 Br_{t+1}}{180}\right), \\ & y_{t} + Le \cdot \sin
 \left(\frac{\pi \cdot Br_{t+1}}{180}\right) \\	
 when  \hspace{0.5cm}     & Br_{t} < 180, Br_{t+1} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t} > 180, Br_{t+1} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $Br_t$ is the bearing at time $t$, $k$ the
concentration parameter from the von Mises distribution that we correlate with
the revenue so that $k = (Rev + 1 / RefRev) \cdot max_{k}$, where $max_{k}$ is
the maximum concentration value, $k$, and $RefRev$ is parametrised as for
$\beta_{3}$ in the step length function. Details of the variables, meaning and
units for fleet dynamics are provided in Table \ref{tab:fleet_dyn}.

\subsubsection{Local population depletion}

Where several fishing vessels exploit the same fish population competition is
known to play an important role in local distribution of fishing effort
\citep{Gillis1998}. If several vessels are fishing on the same patch of fish,
local depletion and interference competition will affect fishing location
choice of the fleet as a whole \citep{Rijnsdorp2000, Poos2007a}.  To account
for this behaviour, the fishing sub-model operates spatially on a daily
time-step so that for future days the biomass available to the fishery is
reduced in the areas fished. The cumulative effect is to make heavily fished
areas less attractive as a future fishing location choice as reduced catch
rates will be experienced.

\subsection{Fisheries-independent survey}

A fisheries-independent survey is simulated where fishing on a regular grid
begins each year at the same time for a given number of stations (a fixed
station survey design). Catches of the populations at each station are recorded
but not removed from the population (catches are assumed to have negligible
impact on population dynamics). This provides a fishery independent snapshot of
the populations at a regular spatial intervals each year, similar to scientific
surveys undertaken by fisheries research agencies. \\

\subsection{Software: R-package development}

The simulation framework is implemented in the statistical software package R
\citep{RCoreTeam2017} and available as an R package from the author's github
site (\url{www.github.com/pdolder/MixFishSim}).\\

\section{Model calibration}
We calibrate \textit{MixFishSim} to investigate the influence of data
aggregation on spatial inference.

\subsection{Population models}

We calibrated the simulation model for four example populations with different
demographics, growth rates, natural mortality and recruitment (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
(Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4}) defined to be
unique to each population resulting in differently weekly distribution patterns
(Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was
assumed to have two defined spawning areas that result in the populations
moving towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}).  \added{The population
demographics were chosen to broadly represent three mobile low-medium value
groundfish species and one high value species with low mobility, with the
dynamics hypothetical but \deleted{as you} might \added{be} expect\added{ed} \deleted{to find }in a typical demersal
fishery.}

\subsection{Fleet calibration}

\replaced{F}{The f}leets were calibrated to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}). By setting different
catchability coefficients ($Q_{fl, p}$) we create different targeting
preferences between the fleets and hence \added{different} spatial dynamics.  The learned random
walk process implies that within a fleet different vessels have different
spatial distributions based on individual experience. The step function was
calibrated dynamically within the simulations as the maximum revenue
obtainable was not known beforehand. This was implemented so that vessels take
smaller steps when fishing at a location that yields landings value in the top
90th percentile of the value experienced in that year so far (as defined per
fleet in Table \ref{tab:2}). \\

Fishing locations were chosen based on random search and, with increasing
proportion as time progressed, experience of profitable catches built up in the
same month from previous years and from the previous trip.  `Profitable' in
this context was defined as the locations where the top 70 \% of expected
profit would be found given revenue from previous trips and cost of movement to
the new fishing location. This probability was based on a logistic sigmoid
function with a lower asymptote of 0 and upper asymptote of 0.95, and a \replaced{slope}{growth
rate} that ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

\subsection{Survey settings}

The survey simulation was set up with a fixed gridded station design with 100
stations fished each year, starting on day 92 and ending on day 112 (5 stations
per day) with same catchability parameter ($Q_{p} = 1$) for all populations\deleted{ $p$}.
This approximates a real world survey design with limited seasonal and spatial
coverage.

\subsection{Example research question}

To illustrate the capabilities of \textit{MixFishSim}, we investigate the
influence of the temporal and spatial resolution of different data sources on
the reduction in catches of a population given spatial closures. To do so, we
set up a simulation to run for 50 years based on a 100 $\times$ 100 square grid
(undetermined units), with five fleets of 20 vessels each and four fish
populations. Fishing takes place four times a day per vessel and five days a
week, while population movement is every week. \\

\textit{How does sampling-derived fisheries data reflect the underlying
	population structure?} \\

To answer this question we compare different spatial and temporal aggregations
of the true population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population
		density from a fixed-site sampling survey design as commonly
		used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population
		density from our fleet model that includes fishery-induced
		sampling dynamics.
\end{enumerate}

We allow the simulation to run unrestricted for 30 years, then implement
spatial closed areas for the last 20 years of the simulation based on data
(either derived from the commercial catches, fisheries-independent survey or
the true population) used at different spatial and temporal scales. \\

The following steps are undertaken to determine closures:

\begin{enumerate}
	\item Extract data source (true population, commercial or survey), 
	\item Aggregate according to desired spatial and temporal resolution,
	\item Interpolate across entire area at desired resolution using simple
		bivariate interpolation using the \emph{interp} function from
		the R package akima \citep{Akima2006}. This is intended to
		represent a naive spatial model of catch rates, without
		knowledge of the spatial population dynamics.
	\item Close area covering top 5 \% of catch rates.
\end{enumerate}
In total 28 closure scenarios were run that represent combinations of:

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		true population,
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid,

\end{itemize}

\added{We implemented a series of spatial closures targeted at reducing fishing
	mortality on population 3, given the different data sources and spatial
	and temporal resolutions above. We use the effectiveness of these
	closures in reducing fishing mortality as a way of evaluating the
	trade-offs in data sources and resolution.} Survey closures were on an
annual basis only, as this was the most temporally resolved survey data
available. We evaluated the factors contributing to the success of the closures
through a regression tree (using the R package REEMtree \citep{Sela2012}) to
identify the factor most contributing to differences in fishing mortality
before and after the closure.

\section{Results}

\subsection{Emergent simulation dynamics}

Individual habitat preferences and thermal tolerances result in different
spatial habitat use for each population (Figure \ref{S-fig:9}) and consequently
different seasonal exploitation patterns (Figure \ref{S-fig:10}).

It can be seen from a single vessels movements during a trip that the vessel
exploits three different fishing grounds, each of them multiple times (Figure
\ref{fig:mv}A), while across several trips fishing grounds that are further
apart are fished (Figure \ref{fig:mv}B). These different locations relate to
areas where the highest revenue were experienced, as shown by Figure
\ref{fig:mv}D, where several vessels tracks are overlaid on the revenue
field.

Vessels from the same fleet (and therefore targeting preference) may exploit
some shared and some different fishing grounds depending on their own personal
experience during the exploratory phase of the fishery (Figure \ref{fig:mv} (C)).
This results from the randomness in the correlated random walk step function,
with distance moved during the exploitation phase and the direction
stochastically related to the revenue experienced on the fishing ground (Figure
\ref{fig:mv} (D)).

\subsection{How does sampling-derived fisheries data reflect the underlying
	population structure?}

\deleted{The }\deleted{aggregated }\replaced{C}{c}atch composition \added{aggregated at different
	spatial resolutions} from each of the data sources (\deleted{which
	shows }average seasonal patterns over a ten-year period) highlights
different patterns in perceived community structure depending on the data
source and aggregation level (Figure \ref{fig:1}). The finer spatial grid for
the true population (top left) and commercial data (top middle) show visually
similar patterns, though there are large unsampled areas in the commercial data
from a lack of fishing activity (particularly in the lower left part of the
sampling domain). \replaced{S}{The s}urvey data at this spatial resolution displays very
sparse information about the spatial distributions of the populations. The
slightly aggregated data on a 5 x 5 grid shows similar patterns and, while
losing some of the spatial detail, there remains good consistency between the
true population and the commercial data. Survey data starts to pick out some of
the similar patterns as the other data sources, but lacks spatiotemporal
coverage. The spatial catch information on a 10 x 10 and 20 x 20 grid lose a
significant amount of information about the spatial resolutions for all data
sources, and some differences between the survey, commercial and true
population data emerge. \\

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data at different timescales, with weekly (top), monthly
(middle) and yearly (bottom) catch compositions from across an aggregated 20 x
20 area showing different patterns (Figure \ref{fig:2}). In the true
population, the monthly aggregation captures the major patterns of composition
seen in the weekly data with the percentage of different populations in the
catch having similar mean and standard deviations (Table \ref{tab:7}). In the
weekly and monthly data population 2 dominates. However, some of the variation
was lost when aggregated to an annual level, as indicated from the lower
standard deviations (Table \ref{tab:7}).\\

\replaced{Weekly}{The} commercial data \deleted{on a weekly basis }shows some
of the same patterns as the true population, though \deleted{the }population 1
\deleted{(in red)} is less well represented and some weeks are missing catches
from the area. Here, weekly and monthly compositions were nearly identical
(Figure \ref{fig:2}; Table \ref{tab:7}). \deleted{Again, y}\added{Y}early
values h\deleted{e}ad a similar mean but smaller standard deviation. The survey
data was only available on an annual basis, and showed again a slightly
different composition from the true population and the commercial data; in
particular a greater proportion of population 4 (Figure \ref{fig:2}).

\subsection{How does data aggregation and source impact on spatial fisheries
	management measures?}

\deleted{We implemented a spatial closure using the different data sources and
	spatial and temporal aggregations as outlined in the protocol in
	Section 3.4. We used this to assess the efficacy of a closure in
	reducing fishing mortality on population 3, given availability of data
	and its use at different resolutions in order to evaluate the
	trade-offs in data sources. }

In most cases the fishery closure was successful in reducing fishing mortality
on the species of interest (population 3; Figure \ref{fig:3}). Interestingly
the largest reductions in fishing mortality happened immediately after the
closures, following which the fisheries ``adapted" to the closures \added{by
	finding new areas of high abundance to fish. This led to} fishing
mortality increasing again, though not to past levels (Figure \ref{fig:3}).
The exception to the success was the closures implemented based on the coarsest
spatial (20 x 20) and temporal resolution (yearly) that was ineffective (i.e.
failed to reduce fishing mortality) with all data sources. As expected,
closures based on the ``known" population distribution were most effective,
with differing degrees of success using the commercial data. Fishing mortality
rates on the other species changed in different proportions, depending on
whether the displaced fishing effort moved to areas where the populations were
found in greater or lesser density. \\

The factor most contributing to differences in fishing mortality before and
after the closure was the population (72 \% showing that the closures were
effective for population 3), followed by \added{spatial} data resolution (21 \%), data type (7
\%) with the least important factor the timescale ($<$ 1 \%). In general the
finer the spatial resolution of the data used the greater reduction in fishing
mortality for population 3 after the closures (Figure \ref{fig:4}). The notable
outliers are the commercial data at the coarsest spatial resolution (20 x 20)
at a yearly and weekly timescale, where closures were nearly as effective as
the fine-scale resolution. In this case the closures were sufficiently large to
protect a core area of the habitat for the population, but this was achieved in
a fairly crude manner by closing a large area - including area where the
species was not found (Figure \ref{fig:17}) that may have consequences in terms
of restricting the fishery in a much larger area than necessary. We found that
these trade-offs existed, with high catches maintained with an effective
closure when the highest resolution data was used, with the effect being linear
when the true population distribution was known and also persisting for
closures based on commercial information (Figure \ref{fig:ce}).\\

\section{Discussion}

%% General intro
Our study presents a new highly resolved fisheries simulation framework to
evaluate the importance of data scaling and considers potential bias introduced
through data aggregation when using fisheries data to infer spatiotemporal
dynamics of fish populations. Understanding how fishers exploit multiple
heterogeneously distributed fish populations with different catch limits or
conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to species targeting in fisheries resulting in preferential sampling
\citep{Martinez-Minaya2018}. Often data are aggregated or extrapolated which
requires assumptions about the spatial and temporal scale of processes. Our
study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatiotemporal
approaches to maximise productivity while taking account of both the biological
and human processes operating on different time-frames \citep{Dunn2016},
understanding assumptions behind the data used - increasingly a combination of
logbook and positional information from vessel monitoring systems - is vital to
ensure measures are effective. \\

\subsection{Simulation dynamics}

We employ a simulation approach to model each of the population and fishery
dynamics in a hypothetical `mixed fishery', allowing us to i) evaluate the
consequences of different aggregation assumptions on our understanding of the
spatiotemporal distribution of the underlying fish populations, and ii)
evaluate the effectiveness of a spatial closure given those assumptions. \\

%% USP
Our approach is unique in that it captures fine scale population and fishery
dynamics and their interaction in a way not usually possible with real data and
thus not usually considered in fisheries simulations. While other simulation
frameworks seek to model individual vessel dynamics based on inferred dynamics
from VMS and logbook records \citep{Bastardie2010}, or as a system to identify
measures to meet particular management goals \citep{Bailey2018}, our framework
allows users to explore \deleted{the} assumptions in modelling observational
data and \added{to} evaluate the underlying dynamics of such approaches at
\deleted{a} fine spatial and temporal scale\added{s}.  This offers the
advantage that larger scale fishery patterns are emergent properties of the
system and results can be compared to those obtained under a statistical
modelling framework. \\

%% Comparison to other models
Typically, simulation models that treat fish as individuals are focussed on
exploring the inter- and intra- specific interactions among fish populations
(e.g. OSMOSE; \cite{Shin2004}) in order to understand how they vary over space
and time. Our focus was on understanding the strengths and limitations of
inference from catch data obtained through commercial fishing activity with
fleets exploiting multiple fish populations\added{. This shows how}\deleted{
	and} realis\added{ed}\deleted{ing} catch distributions \deleted{that
}may differ from the underlying populations, as identified by
\cite{Gillis2008}. As such, we favoured a minimum realistic model of the fish
populations \citep{Plaganyi2014} taking account of environmental but not
demographic stochasticity, while incorporating detailed fishing dynamics that
take account of different drivers in a mechanistic way.  \\

Demographic stochasticity arises due to individual-level variability in time to
reproduction and death. This form of stochasticity is often modelled by drawing
random time intervals from a given distribution \citep{Gillespie1977}. The
impact of demographic stochasticity depends on the population size, with the
effects expected to decrease with increasing population size \citep{Lande2010}.
This contrasts with environmental stochasticity, which affects all population
sizes and is present at the population level in our model by variability in
recruitment. \\

We take account of heterogeneity in fleet dynamics due to different preferences
and drivers similarly to other approaches \citep{Fulton2011}, but at an
individual vessel rather than fleet level. We do not explicitly define fleets
as rational profit maximisers at the outset, but consider there are several
stages to development of the fishery; information gathering through search
where the resource location is not known, followed by individual learnt
behaviour of profitable locations.  This provides a realistic model of how
fishing patterns are established and maintained to exploit an uncertain
resource through an explore-exploit strategy \citep{Mangel1983, Bailey2018}.

\subsection{How does sampling-derived fisheries data reflect the underlying
	population structure?}

%% Why ours is important

Our results demonstrate the importance of considering data scale and resolution
when using observational data to support management measures. We find that
understanding of the community composition dynamics will depend on the level of
data aggregation and its important to consider the scale of processes;
including population movement rates, habitat uniformity and fishing targeting
practices if potential biases in data are to be understood and taken into
account (Figures \ref{fig:mv} and \ref{S-fig:9}). \\

Our simulation shows that, despite biases introduced through the fishing
process, the commercially derived data could still inform on the key spatial
patterns in the community structures where the fisheries occurred, which was
spatially limited due to the ``hotspots" of commercially valuable species being
fished. Similarly, despite \deleted{the }even spatial coverage the survey \deleted{was able to
}capture\added{d} some of the same spatial patterns as the true population, but missed
others due to gaps between survey stations limiting spatial and temporal
coverage (Figure \ref{fig:1}). This provides a challenge when
modelling unsampled areas in inferring species distribution maps, though these
limitations may be overcome by understanding the relationship between the
species and habitat covariates where these are known at unsampled locations
\citep{Robinson2011}. \\

%% key findings in context
\subsection{How does data aggregation and source impact on spatial fisheries
	management measures?}

From our simulations spatial disaggregation was more important than the
temporal disaggregation of the commercial data. This reflects the fact that
there was greater spatial heterogeneity over the spatial domain than
experienced in \replaced{given}{individual} locations over the course of the year (Figure
\ref{S-fig:9}). \\

The yearly data assumes the same proportion of each population caught at any
time of the year due to the data aggregation. This assumption introduces
`aggregation bias' as the data may only be representative of some point (or no
point) in time. The monthly data shows some consistency between the real
population and commercial data for population 2 - 4, though population 1
remains under-represented.  On an annual basis, interestingly the commercial
data under represents the first species while the survey over represents
species 1. This is likely due to the biases in commercial sampling, with the
fisheries not targeting the areas where population 1 are present and the survey
sampling areas where population 1 is more abundant than on average. This
indicates that fixed closures, at the right resolution, when based on
commercially derived data have the potential to reduce\deleted{d} fishing
mortality. The likely cost of poor spatial and temporal resolution is
associated with reduced effectiveness and potentially closing fishing
opportunities for other fisheries (Figure \ref{fig:ce}). \\

Two contrasting real world approaches in this respect were the spatial closures
to protect cod in the North Sea. In one example, large scale spatial closures
were implemented with little success due to effort displacement to previously
unfished areas \citep{Dinmore2003}, while in another small scale targeted
spatiotemporal closures were considered to have some effect in reducing cod
mortality without having to disrupt other fisheries substantially
\citep{Needle2011}. These examples emphasise the importance of considering the
right scale and aggregation of data when identifying area closures and the need
to consider changing dynamics in the fisheries in response to such closures. \\

Our study showed that fishing rates on other populations also changed (both up
and down) as a side-effect of closures to protect one species.  This indicates
the importance \replaced{of}{in} considering fishing effort reallocation following spatial
closures, and our simulation allows us to consider the spatiotemporal reasons
for these changes.

\subsection{Model assumptions and caveats}

We modelled the population and fleet dynamic processes to draw inference on the
importance of data scale and aggregation in understanding and managing mixed
fisheries and their impact on multiple fish populations. In doing so, we
necessarily had to make a number of simplifying assumptions. \\

Fish populations in our simulations move in pre-defined timescales and
according to fixed habitat preferences and temperature gradients (Figures
\ref{S-fig:1}, \ref{S-fig:3}). Our assumptions in calibrating the model
(movement rates, temperature tolerances) will have a direct impact on our
conclusions on the relative importance of spatial and temporal processes. These
assumptions could be explored in a future study by varying the parameters and
assessing the robustness of our conclusions. For our example application we
have chosen movement rates to reflect aggregation periods observed in past
studies \citep{Poos2007}. \\

In addition, we have assumed that fishing vessels are not restricted by quota
and therefore discarding of species for which vessels have no quota or that are
unwanted is not taken into account. This is likely to be a significant source
of bias in any inference using commercial data and should also be explored. For
example, \textit{MixFishSim} could be altered to allow for spatiotemporal
appraisal of the impact of discarding on fisher behaviour and underlying
populations via inclusion as discarding behaviour, or through move-on rules or
cessation of fishing activity when quota is exhausted. \\

\subsection{Future applications of \textit{MixFishSim}}

We consider that the increased availability of high resolution catch and
locational information from commercial fisheries will \replaced{make}{require} it \deleted{to be }a key
source of data for ensuring management is implemented at the right scale in
future. For example, identifying hot-spots for bycatch reduction or identifying
spatial overlaps in mixed fisheries \citep{Dolder2018, Gardner2008, Little2014,
	Dedman2015, Ward2015}. Our simulation model has the potential to test
some of the assumptions behind the modelling approaches in identifying such
hotspots and indeed behind spatiotemporal modelling in general\replaced{, }{
	(}e.g. comparing GAMs, GLMMs, Random Forests and geostatistical models
under different data generation processes as exampled by
\cite{Stock2019}\deleted{)}. \\

Other novel applications of our framework could be: testing different survey
designs given multiple species and data generating assumptions \citep{Xu2015};
commercial index standardisation methods and approaches and understanding of
appropriate scales and data aggregations and non-proportionality in catch rate
and abundance \citep{Harley2001, Maunder2004}; exploring assumptions about the
distribution of natural mortality and fishing mortality throughout the year and
importance of capturing in-year dynamics in estimating stock status
\citep{Liu2013}; at\added{-}sea sampling scheme designs to deliver unbiased estimates
of population parameters \citep{Cotter2007, Kimura2006}; adaptive management
\citep{Walters2007, Dunn2016}; testing the ability of commonly employed fleet
dynamics models such as Random Utility Models to capture fine scale dynamics
and understand their importance \citep{Girardin2016}; and as a detailed
operating model in a management strategy evaluation \citep{Mahevas2004}. \\

\section{Conclusions}

\textit{MixFishSim} provides a detailed simulation framework to explore the
interaction of multiple fisheries exploiting different fish populations. The
framework enables users to evaluate assumptions in modelling commercially
derived data through comparison to the true underlying dynamics at a fine
spatial and temporal scale.  Understanding these dynamics, the limitations of
the data and any potential biases that may be introduced when making inference
on spatiotemporal interactions will enable users to identify weaknesses in
modelling approaches and identity where data collection is needed to strengthen
inference. \\

Our application shows that inference on community dynamics may change depending
on the scale of data aggregation. There is an important balance in ensuring
that the data are sufficiently spatially and temporally disaggregated that the
main features of the data are captured, yet maintaining enough data coverage
that the features can be distinguished. We found \deleted{in our application
}greater spatial than temporal heterogeneity. When using aggregated data to
define spatial closures coarser temporal resolution (months instead of weeks)
could still achieve the same results in reducing exploitation rates of a
vulnerable species at the highest temporal resolution data. Conversely,
reducing the spatial resolution had a negative effect on the effectiveness of
the measures \deleted{(}though\added{,} importantly, there was still some
benefit even with coarse spatial resolution\deleted{)}. \\

While \deleted{any findings are likely to be} case\added{-}specific, our
findings emphasise the need to understand population demographics, habitat use
and movement rates in designing any closure scenario based on observational
sampling. This information can then be used to set the bounds on data
aggregation used in modelling studies aimed at informing the management
measures. \\

\deleted{\textit{MixFishSim} has numerous potential additional applications as it
enables the user to apply methods to a fisheries system where there is detailed
understanding of underlying spatiotemporal dynamics. This enables
identification of weaknesses or limitations which would not be possible
otherwise. In future, we recommend use of the framework to test
hypothes\replaced{e}{i}s that are otherwise unable to be analysed using real
world data due to limitations of data collection. That way the knowledge gained
through simulation can inform the future design of management measures.}

\section*{Funding and Acknowledgements} This work was supported by the MARES
doctoral training program (MARES\_14\_15) and the Centre for Environment,
Fisheries and Aquaculture Science seedcorn program (DP227AC).\added{ We thank
	two anonymous reviewers for their helpful comments which greatly
	improved the final manuscript.} The authors declare no competing
interests.

\newpage

\input{parameter_tables.tex}

\clearpage
\input{F_changes.tex}

\clearpage
\input{Temp_proportions.tex}

%%%%%%%%%%%%%%%%%
%%%% figures %%%%
%%%%%%%%%%%%%%%%%

%% the tikz figure
\clearpage
\input{Process_model.tex}
%\begin{comment}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Combined_Movement}
	\caption{(A) The fishing locations (points) and movements (lines) of a
		single vessel during a trip overlaid on the revenue of a
		fishing site (landings x price; darker purple = higher
		revenue); (B) the fishing locations of the vessel over several
		trips (value field changes over the period so not shown). Note
		that movements are a mixture of correlated random walk (solid
		lines) and experience-based (dashed lines), and that the field
		is wrapped on a torus so that opposite sides of the spatial
		domain are considered spatially close; (C) the locations of
		multiple vessels from the same fleet overlaid on the value
		field, (D) the realised step distance and turning angles for a
		single vessel over the simulation.}
	\label{fig:mv}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{[Colour] Data aggregation at different spatial resolutions
		over a ten year period. \added{The figure shows catch
			composition at each spatial unit represented by a
			square pie chart of the four populations. The area of
			each colour is proportional to the weight of each
			population caught in that unit. Figure produced using
			the R package `mapplots' (\cite{Gerritsen2014}.}
	}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_time_Rev2}
	\caption{[Colour]Proportion of each population (y axis) for data aggregated at
		different temporal resolutions. Data is aggregated over a
		ten-year period for an area 20 x 20. Each bar represents either
		a week, month or year respectively.}
	\label{fig:2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/F_trendsREV}
	\caption{[Colour]Comparison of closure scenarios effect on fishing mortality
		trends. Line colour denotes timescale, while linestyle
		denotes spatial resolution. The vertical dashed line
		indicates the onset of the spatial closures.}
	\label{fig:3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{./Plots/f_diff_effectiveness}
	\caption{Comparison of closure scenario effectiveness based on
		different spatial and temporal resolutions.}
	\label{fig:4}
\end{figure}

\begin{figure}[!ht]
	\centering
	\includegraphics[width =0.5\linewidth]{./Plots/Closure_fishing_locations_yearly}
	\caption{The location of fishing effort, (a) before the spatial closure
		and (b) after the spatial closure (years in panel), and (c) the
		suitable habitat for population 3. The site of the closure
		can be seen in the red box on all three panels.}
	\label{fig:17}
\end{figure}

\begin{figure}[!ht]
	\centering
	\includegraphics[width =\linewidth]{./Plots/Closure_effectiveness_CF}
	\caption{[Colour]Effectiveness of closure with regards to reducing fishing
		mortality on protected population (further left on x-axis is
		best) and maintaining high catches in the fishery (highest on
		y-axis is best). The numbers indicate the spatial resolution of
		the data, while grey lines indicate the direction of the
		trade-off between reducing fishing mortality and overall
		catches.}
	\label{fig:ce}
\end{figure}

%\end{comment}

\clearpage

\section*{References}

\bibliography{simulation_framework}

\end{document}